---
title: "Calculation"
output: html_document
---

Questions to answer:

##1. 
Compute analytically the probability that the Braves win the world series when the sequence of game locations is {NYC, NYC, ATL, ATL, ATL, NYC, NYC}. (The code below computes the probability for the alternative sequence of game locations. Note: The code uses data.table syntax, which may be new to you. This is intential, as a gentle way to introduce data.table.) Calculate the probability with and without home field advantage when PB = 0.55. What is the difference in probabilities?

###With home field advantage
```{r}
require(dplyr)
require(data.table)
# Get all possible outcomes
apo <- fread("deliverables/assets/all-possible-world-series-outcomes.csv")

# Home field indicator
#hfi <- c(1,1,0,0,0,1,1) #{ATL, ATL, NYC, NYC, NYC, ATL, ATL}
hfi <- c(0,0,1,1,1,0,0)

# P_B
pb <- 0.55
advantage_multiplier <- 1.1 # Set = 1 for no advantage
pbh <- 0.55*advantage_multiplier
pba <- 1 - (1 - 0.55)*advantage_multiplier
```

```{r, cache = TRUE}
# Calculate the probability of each possible outcome
apo[, p := NA_real_] # Initialize new column in apo to store prob
for(i in 1:nrow(apo)){
  prob_game <- rep(1, 7)
  for(j in 1:7){
    p_win <- ifelse(hfi[j], pbh, pba)
    prob_game[j] <- case_when(
        apo[i,j,with=FALSE] == "W" ~ p_win
      , apo[i,j,with=FALSE] == "L" ~ 1 - p_win
      , TRUE ~ 1
    )
  }
  apo[i, p := prod(prob_game)] # Data.table syntax
}

# Sanity check: does sum(p) == 1?
apo[, sum(p)] # This is data.table notation

# Probability of overall World Series outcomes
apo[, sum(p), overall_outcome]
```
With home field advantage:
P(win WS) = 0.604
p(lose) = 0.396

This is problematic!!
```{r}
p_lose <- 0
for(k in 0:3){
  p_lose <- p_lose + pbinom(4-k, 4, pba) * dbinom(k, 3, pbh)
  print(p_lose)
}
```

### Without home field advantage
```{r}
# Calculate the probability of each possible outcome
apo[, p := NA_real_] # Initialize new column in apo to store prob
for(i in 1:nrow(apo)){
  prob_game <- rep(1, 7)
  for(j in 1:7){
    p_win <- ifelse(hfi[j], pb, pb)
    prob_game[j] <- case_when(
        apo[i,j,with=FALSE] == "W" ~ p_win
      , apo[i,j,with=FALSE] == "L" ~ 1 - p_win
      , TRUE ~ 1
    )
  }
  apo[i, p := prod(prob_game)] # Data.table syntax
}

# Sanity check: does sum(p) == 1?
apo[, sum(p)] # This is data.table notation

# Probability of overall World Series outcomes
apo[, sum(p), overall_outcome]
```
Without home field advantage, P(win WS) = 0.608

### Difference
```{r}
diff_theory <- abs(0.604221 - 0.6082878)
```


## 2. Calculate the same probabilities as the previous question by simulation.

### Without home advantage
```{r, cache =  TRUE}
N <- 10000          # number of repetition in the simulation
outcome <- rbinom(N, 7, pb)
p_win_no_advantage <- mean(outcome >= 4)
```
P(win WS | no home field advantage) = 0.6072


### With home advantage
```{r, cache = TRUE}
hfi <- c(0,0,1,1,1,0,0)
p_win <- rep(NA, 7)
for(j in 1:7){
    p_win[j] <- ifelse(hfi[j], pbh, pba)
}
```

```{r, cache= TRUE}
N<-100000
win_WS <- rep(NA, N)

for(k in seq(win_WS)){
  sum_one_series <- 0
  for (i in 1:7){
    sum_one_series <- sum_one_series + rbinom(1,1, p_win[i])
    #print(sum_one_game)
    if(sum_one_series >= 4){
      win_WS[k] = 1
    }
    else{
      win_WS[k] = 0
    }
  }
}
p_win_with_advantage <- mean(win_WS)
```
P(win WS | home field advantage) = 0.606

### Difference
```{r}
diff_simulation <- abs(p_win_with_advantage - p_win_no_advantage)
```

## Problem 3
3. What is the absolute and relative error for your simulation in the previous question?

```{r}
abs_error <- abs(diff_theory - diff_simulation)
rel_error <- abs_error/diff_theory
```


4. Bonus. Does the difference in probabilites (with vs without home field advantage) depend on PB?



5. Bonus. Does the difference in probabilites (with vs without home field advantage) depend on the advantage factor? (The advantage factor in PBH and PBA is the 1.1 multiplier that results in a 10% increase for the home team.)