---
title: "calculation"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
require(stats4)
```

Step 3: Calculate a 95% confidence interval from the approximated sampling distribution. Given the sample data from step 2, we generate a 95% confidence interval by using the quantile function to capture the middle 95% data.

```{r}
generate_ci <- function(data, alpha = 0.1){
  quantile(data, c(alpha/2, 1 - alpha/2))
}
```

Step 4: Examine the definition of coverage probability in part I. Here, for each confidence interval generated from samples, we check if it covers the underlying true value. Then we calculate the percentage of if successfully covering the null value, and get our coverage probability.

```{r}
capture_result <- function(ci, par){
  1*(ci[[1]] < par & par < ci[[2]])
}
```



# Part I: Generate sample

```{r}
N <- 201

generate_data <- function(N, dis){
  if (dis == "norm"){
    sample <- rnorm(N, mean = 0, sd = 1)
  } else if (dis == "gamma") {
    sample <- rgamma(N, shape = 1.4, scale = 3)
  }
  sample
}
```

# Part II: Models
```{r}
# Method of moments with normal
# Generate Distributions
# par: Parameter of interest (median or min)
MME_normal <- function(sample, par, R = 100, N = 201){
  xbar <- mean(sample)
  s2 <- var(sample)
  
  mean_hat <- xbar
  mean_sd <- sqrt(s2)
  
  d1 <- rnorm(N*R, mean_hat, mean_sd) %>% array(dim = c(R, N))
  d2 <- d1 %>% apply(1, par)
  d2
}
```


```{r}
# Method of moments with gamma
# Generate Distributions
# par: Parameter of interest (median or min)
MME_gamma <- function(sample, par, R = 100, N = 201){
  xbar <- mean(sample)
  s2 <- var(sample)
  
  shape_hat <- xbar^2/s2
  scale_hat <- s2/abs(xbar)
  
  d1 <- rgamma(N*R, shape = shape_hat, scale = scale_hat) %>% array(dim = c(R, N))
  d2 <- d1 %>% apply(1, par)
  d2
}
```

```{r}
# Kernal Density Estimation
# Generate Distributions
# par: Parameter of interest (median or min)
KDE <- function(sample, par, R = 100){
    ecdfstar <- function(t,sample,smooth){
      outer(t,sample,function(a,b){pnorm(a,b,smooth)}) %>% rowMeans
    }
    
    tbl <-data.frame(
      x = seq(min(sample)-2*sd(sample),max(sample)+2*sd(sample),by=0.01)
    )
    
    tbl$p <-ecdfstar(tbl$x,sample,0.3)
    tbl <- tbl[!duplicated(tbl$p),]
    
    qkde <- function(ps,tbl){
      rows <- cut(ps,tbl$p,labels=FALSE)
      tbl[rows,"x"]
    }
    
    U <- runif(N*R)
    d1 <- array(qkde(U,tbl),dim=c(N,R))
    d2 <- apply(d1,2,par)
    d2
}
```

```{r}
# Bootstrap
# Generate Distributions
# par: Parameter of interest (median or min)
bootstrap <- function(sample, par, R = 100, N = 201){
  func <- get(par)
  d1 <- rep(NA, R)
  for(i in 1:R){
    index <- sample.int(N, N, replace = TRUE)
    d1[i] <- sample[index] %>% func(na.rm = TRUE)
  }
  d1
}
```


# Part IV: Simulation
```{r, cache = TRUE}
# simulation
N <- 201
M <- 1000
R <- 1000 

distribution <- c("norm", "gamma")
method <- c("MME_normal", "MME_gamma", "KDE", "bootstrap")
parameter <- c("median", "min")

sim_par <- c()
for(dis in distribution){
  for(par in parameter){
    out <- rep(NA, M)
    for(i in 1:M){
        out[i]<- N %>% generate_data(dis) %>% get(par)()
    }
  sim_par <- append(sim_par, mean(out))
  }
}

```

```{r}
M <- 50
for(dis in distribution){
  for(m in method){
    func <- get(m)
    for(par in parameter){
      if(dis == "norm" & par == "median"){
        true_par <- 0
      }else if (dis == "norm" & par == "min"){
        true_par <- sim_par[2]
      }else if (dis == "gamma" & par == "median"){
        true_par <- sim_par[3]
      }else if (dis == "gamma" & par == "min"){
        true_par <- sim_par[4]
      }
      capture <- rep(NA, M)
      for (i in 1:M){
        
       sample_dis  <- N %>% generate_data(dis) %>% func(par)
       capture[i] <- sample_dis %>% generate_ci(0.1) %>% capture_result(true_par)
      }
      mean <- mean(capture)
      print(paste(dis, m, par, mean))
    }
  }
}
```

