---
title: "Simulation Study"
author: "Yan Shen"
date: 11/03/2019
output: 
  html_document:
    #code_folding: hide
    theme: journal
    toc: yes
    toc_depth: 5
    toc_float: true
editor_options: 
  chunk_output_type: inline
---

***
# I. Introduction

  In this blog post, we want to study how different methods affect coverage probability. 
  
> __Coverage probability__ is the long run proportion of intervals that capture the population parameter of interest. 

  Idealy, a 95% confidence interval will capture the population parameter of interest in 95% of samples.
  
***
# II. Setup

  Here, we use various methods to calculate the __90%__ confidence interval by performing a __$2*4*2$ factorial simulations__. We use a sample size of __N = 201__ through the study.
  
  The three variables are: 
  
  + True, underlying distribution
    1. Standard normal distribution
    2. Gamma distribution (shape = 1.4, scale = 3)
  + Model
    1. Method of moments with normal distribution
    2. Method of moments with gamma distribution
    3. Kernel Density Estimation
    4. Bootstrap
  + Parameter of interest
    1. Sample min
    2. Median
  
***
# III. Functions

## Part I: Generate sample

  Set N = 201. Then generate sample according to the parameter, "norm" for standard normal distribution and "gamma" for gamma distribution.
```{r}
N <- 201

generate_data <- function(N, dis){
  if (dis == "norm"){
    sample <- rnorm(N, mean = 0, sd = 1)
  } else if (dis == "gamma") {
    sample <- rgamma(N, shape = 1.4, scale = 3)
  }
  sample
}
```

## Part II: Models

  For different models, we __generate a distribution__ for the parameter we are interested in. The parameter can be either __median or min__ in this study.
  
### 1. Method of moments with normal
```{r}
# Method of moments with normal
# Generate Distributions
# par: Parameter of interest (median or min)
MME_normal <- function(sample, par, R = 1000, N = 201){
  xbar <- mean(sample)
  s2 <- var(sample)
  
  mean_hat <- xbar
  mean_sd <- sqrt(s2)
  
  d1 <- rnorm(N*R, mean_hat, mean_sd) %>% array(dim = c(R, N))
  d2 <- d1 %>% apply(1, par)
  d2
}
```

### 2. Method of moments with gamma
```{r}
# Method of moments with gamma
# Generate Distributions
# par: Parameter of interest (median or min)
MME_gamma <- function(sample, par, R = 1000, N = 201){
  xbar <- mean(sample)
  s2 <- var(sample)
  
  shape_hat <- xbar^2/s2
  scale_hat <- s2/abs(xbar)
  
  d1 <- rgamma(N*R, shape = shape_hat, scale = scale_hat) %>% array(dim = c(R, N))
  d2 <- d1 %>% apply(1, par)
  d2
}
```

### 3. Kernal Density Estimation
```{r}
# Kernal Density Estimation
# Generate Distributions
# par: Parameter of interest (median or min)
KDE <- function(sample, par, R = 1000, N = 201){
    ecdfstar <- function(t,sample,smooth){
      outer(t,sample,function(a,b){pnorm(a,b,smooth)}) %>% rowMeans
    }
    
    tbl <-data.frame(
      x = seq(min(sample)-2*sd(sample),max(sample)+2*sd(sample),by=0.01)
    )
    
    tbl$p <-ecdfstar(tbl$x,sample,0.3)
    tbl <- tbl[!duplicated(tbl$p),]
    
    qkde <- function(ps,tbl){
      rows <- cut(ps,tbl$p,labels=FALSE)
      tbl[rows,"x"]
    }
    
    U <- runif(N*R)
    d1 <- array(qkde(U,tbl),dim=c(N,R))
    d2 <- apply(d1,2,par)
    d2
}
```


### 4. Bootstrap
```{r}
# Bootstrap
# Generate Distributions
# par: Parameter of interest (median or min)
bootstrap <- function(sample, par, R = 1000, N = 201){
  func <- get(par)
  d1 <- rep(NA, R)
  for(i in 1:R){
    index <- sample.int(N, N, replace = TRUE)
    d1[i] <- sample[index] %>% func(na.rm = TRUE)
  }
  d1
}
```

## Part III. Helper functions

  We use helper functions to generate confidence interval for the given sampling distribution and see if the confidence interval captures the true underlying parameter.
```{r}
generate_ci <- function(data, alpha = 0.1){
  quantile(data, c(alpha/2, 1 - alpha/2))
}
```

```{r}
capture_result <- function(ci, par){
  1*(ci[[1]] < par & par < ci[[2]])
}
```

***
# IV. Simulation

  First, we simulate the true underlying parameter of interest using simulation. 

```{r, cache = TRUE}
require(magrittr)
# simulation
N <- 201
M <- 5000
R <- 1000 

distribution <- c("norm", "gamma")
method <- c("MME_normal", "MME_gamma", "KDE", "bootstrap")
parameter <- c("median", "min")

sim_par <- c()
for(dis in distribution){
  for(par in parameter){
    out <- rep(NA, M)
    for(i in 1:M){
        out[i]<- N %>% generate_data(dis) %>% get(par)()
    }
  sim_par <- append(sim_par, mean(out))
  print(paste(dis, par, mean(out)))
  }
}
```

  We can see the approximated results:
  
  True Parameter | median | min
  ---------------|--------|-----
   Normal        |   0    |-2.75
   Gamma         |   3.26 | 0.07
  
  Then, for each combination of variables, we generate the coverage probability. 
```{r, cache = TRUE}
M <- 1000

require(foreach)
require(doParallel)

cores_2_use <- detectCores(TRUE) -1
cl <- makeCluster(cores_2_use)
clusterSetRNGStream(cl, 2344)
registerDoParallel(cl)

for(dis in distribution){
  for(m in method){
    func <- get(m)
    for(par in parameter){
      if(dis == "norm" & par == "median"){
        true_par <- 0
      }else if (dis == "norm" & par == "min"){
        true_par <- sim_par[2]
      }else if (dis == "gamma" & par == "median"){
        true_par <- sim_par[3]
      }else if (dis == "gamma" & par == "min"){
        true_par <- sim_par[4]
      }
      capture <- rep(NA, M)
      capture <-  foreach(i = 1:M, .combine = c, .packages = c('dplyr', 'stats4')) %dopar% {
         sample_dis  <- N %>% generate_data(dis) %>% func(par)
         capture[i] <- sample_dis %>% generate_ci(0.1) %>% capture_result(true_par)
      }
      mean <- mean(capture)
      print(paste(dis, m, par, mean))
    }
  }
}
stopCluster(cl)
```

  We can see the approximated results for coverage probability:

  For __normal distribution__: 
  
  Model\\Parameter             | Median | Min
  -----------------------------|--------|--------
  Method of moments with normal|  0.961 | 1
  Method of moments with gamma |  0     | 0
  Kernel Density Estimation    |  0.937 | 0.954
  Bootstrap                    |  0.89  | 0.443
  
  For __gamma distribution__: 
  
  Model\\Parameter             | Median | Min
  -----------------------------|--------|--------
  Method of moments with normal|  0.027 | 0
  Method of moments with gamma |  0.941 | 0.995
  Kernel Density Estimation    |  0.898 | 0.298
  Bootstrap                    |  0.903 | 0.55
  
***
# V. Interpretation

  For both distributions, method of moments with their distribution have the highest coverage probability, while method of moments of the other distribution have the lowest coverage probability. Specifically, for normal distribution, if we use method of moments with gamma, the coverage probability for both median and min are 0. Vice versa, for gamma distribution, the coverage probabilities are close to 0, too. Kernel Density Estimation and Bootstrap approach have similar coverage probabilities.
