---
title: "calculation"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Step1: Generate a single sample from a standard normal distribution of size N = 201. Explain to the reader how you use MLE to estimate the distribution.

```{r}
require(stats4)
# Generate sample
N <- 201
generate_data <- function(N){
  s <- rnorm(N, mean = 0, sd = 1)
  
  # estimate distribution with MLE
  nLL <- function(mean, sd){
    fs <- dnorm(x = s, mean = mean, sd = sd, log = TRUE)
    -sum(fs)
  }
  
  fit <- mle(nLL, start = list(mean = 1, sd = 0), method = "L-BFGS-B", lower = c(0, 0.01))
  
  coef(fit)
}
```


Step2: Show the reader how you approximate the sampling distribution of the median, conditional on the estimate of the distribution in the previous step.

```{r}
# Approximate the sampling distribution of median
generate_distribution <- function(coef, R = 5000){
  out <- rep(NA, R)
  for(i in 1:R){
    b <- rnorm(N, mean = coef[1], sd = coef[2])
    out[i] <- median(b)
  }
  out
}
# hist(out, freq = FALSE, xlab = "sample median")
```


Step3: Describe how you calculate a 95% confidence interval from the approximated sampling distribution.

```{r}
alpha <- 0.05
generate_ci <- function(data, alpha = 0.05){
  quantile(data, c(alpha/2, 1 - alpha/2))
}
```


Step4: Explain the concept of coverage probability. Explain your code for calculating the coverage probability.

```{r}
capture_median <- function(ci){
  1*(ci[[1]] < 0 & 0 < ci[[2]])
}
```


Step5: Perform the simulation and report the results.

```{r}
M <- 10000
captures <- rep(NA, M)

#require(foreach)
#require(doParallel)

cores_2_use <- detectCores(TRUE) -1
cl <- makeCluster(cores_2_use)
clusterSetRNGStream(cl, 2344)
registerDoParallel(cl)

captures <-  foreach(i = 1:M, .combine = c, .packages = c('dplyr', 'stats4')) %dopar% {
  captures[i] <- generate_data(201) %>% generate_distribution %>% generate_ci %>% capture_median
}

stopCluster(cl)

mean(captures)
```


Step6: Describe how you might change the simulation to learn more about the operating characteristics of your chosen method for constructing the 95% confidence interval.

